---
title: "problem_set_8"
author: "Lydia Holley"
date: "2024-03-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Part 1 - Data formatting & Packages

```{r part1}
library(fitdistrplus)

my_file <-read.delim("GSE37642-GPL97_series_matrix.EDITED.txt")
my_data <- data.frame(my_file[,-1], row.names=my_file[,1])

```

# Part 2 - Testing Gene 200000_s_at in Row 1

```{r part2}

row_1 <- my_data[1,]
row_1 <- as.numeric(unlist(row_1))

norm<-suppressWarnings(fitdist(row_1, "norm"))
lnorm<-suppressWarnings(fitdist(row_1, "lnorm"))
gamma<-suppressWarnings(fitdist(row_1, "gamma"))

denscomp(list(norm,lnorm,gamma))

gofstat(list(norm,lnorm,gamma))

```

## Question 1 - Which model has the best fit?\
- the normal model has the best fit because it has the smalled AIC and BIC values

# Part 3 - Testing the first 500 genes for best fit distribution

```{r part3}

num_rows<-1:500
best_models<-list()

for (i in num_rows){
  row_data<-as.numeric(unlist(my_data[i,]))
  
  norm<-suppressWarnings(fitdist(row_data, "norm"))
  lnorm<-suppressWarnings(fitdist(row_data, "lnorm"))
  gamma<-suppressWarnings(fitdist(row_data, "gamma"))
  
  results<-gofstat(list(norm,lnorm,gamma))
  minimum<-min(results$aic)
  model<-match(minimum,results$aic)
  
  best_models<-append(best_models,model)
}

# normal models
num_norm<-length(which(best_models == 1))
pcnt_norm<-num_norm/500

# log normal models
num_lnorm<-length(which(best_models == 2))
pcnt_lnorm<-num_lnorm/500

# gamma models
num_gamma<-length(which(best_models == 3))
pcnt_gamma<-num_gamma/500

```

## Question 2 - Report the percent of the 500 genes that fit the normal, log-normal, and gamma distributions.\
- normal: 0.562
- log normal: 0.394
- gamma: 0.044

# Part 4 - generate the AIC value for gene 200099_s_at (row 100) fit to the above Cauchy distribution

```{r part4}

row_100<-my_data[100,]
num_cols<-1:422
total_log<-0

for (i in 1:length(row_100)){
  value<-as.numeric(unlist(row_100[i]))
  total_log<-total_log + dcauchy(value, location = 14.23807615, scale = 0.07723034, log = TRUE)
  
}
norm_100<-suppressWarnings(fitdist(as.numeric(unlist(my_data[100,])), "norm"))
norm_100$aic

# AIC = -2ln(L) + 2p
aic_cauchy<- (-2*total_log+(2*2))
aic_cauchy

```

## Question 3 - Based on this, does the Cauchy distribution fit better than the normal distribution?\
- the AIC for the normal model is -252.9705 and for the cauchy model is -324.0832. This indicates that the Cauchy model is the better fit.
