---
title: "problem_set_4"
author: "Lydia Holley"
date: "2024-02-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(outliers)

```

# Read in the data

```{r load data}
ratio_data<-read.csv("LSDS-13_calcium-uptake_fajardoTRANSFORMED_1_72_85_93.csv", sep = "\t")
full_data<-read.csv("OSD-488-samples.csv", sep = "\t")

```

# F test

## Question 1

Use the F-test to test equal variances at time-point twenty (20) between\

-   Space_Flight vs Ground Control\
-   F = 1.4396, there is not a statistically significant difference in variances\
-   Space_Flight vs Cohort_Control\
-   F = 2.086, there is not a statistically significant difference in variances\

Report the F-value and if there is a statistically significant difference in the variances

```{r Ftest}

# Space Flight v Ground Control
space_flight<-ratio_data[ratio_data$Spaceflight %in% c("Space_Flight"), ]
space_flight_f<-space_flight$X20
sd(space_flight_f) # has bigger variance, so will go in numerator

ground_control<-ratio_data[ratio_data$Spaceflight %in% c("Ground_Control"), ]
ground_control_f<-ground_control$X20
sd(ground_control_f)

var.test(space_flight_f,ground_control_f)

# Space Flight v Cohort Control

cohort_control<-ratio_data[ratio_data$Spaceflight %in% c("Cohort_Control"), ]
cohort_control_f<-cohort_control$X20
sd(cohort_control_f)

var.test(space_flight_f,cohort_control_f)

```

# T-test - One Sample

## Question 2

At which time point(s) does the mean of all of the tibialis samples not deviate significantly from 0.8?\
- The tibialis samples do not deviate significantly from 0.8 at X280 and X300.

```{r Ttest-onesample}

tibialis<-ratio_data[ratio_data$Tissue %in% c("Tibialis_Anterior"), ]
numSeq<-seq(from = 20, to = 500, by = 20)
fullSeq <- lapply(numSeq, function(x) ifelse(!is.na(x), paste0("X", x), x))
q2_list<-list()

for (i in fullSeq) {
  column<-tibialis[, (names(ratio_data) %in% i)]
  column<-na.omit(column)
  t_test<-t.test(column, mu = 0.8)
  p_value<-as.numeric(t_test$p.value)
  if (p_value>0.05){
    q2_list<-c(q2_list,i)
  }
}

print(q2_list)
```

# T-Test - Two sample Tibialis

## Question 3.1

First, use the F-Test to determine if the variances of the two samples at time points 20 and 480 are statistically different from 1. (not sure if the number 1 is supposed to be the alternative or if I'm supposed to compare the two samples, as in the ratio between tham is 1 or they are equal. for the next question comparison makes the most sense, so that's what I'll do, but I'm confused by the question)\
- the sample variances at X20 and X480 are not significantly different from each other

```{r q3.1}

tibialis<-ratio_data[ratio_data$Tissue %in% c("Tibialis_Anterior"), ]

# Space Flight v Ground Control at X20
space_flight<-tibialis[tibialis$Spaceflight %in% c("Space_Flight"), ]
space_flight_t20<-space_flight$X20
sd(space_flight_t20) 

ground_control<-tibialis[tibialis$Spaceflight %in% c("Ground_Control"), ]
ground_control_t20<-ground_control$X20
sd(ground_control_t20) # has bigger variance, so will go in numerator

var.test(ground_control_t20,space_flight_t20)

# Space Flight v Ground Control at X480
space_flight<-tibialis[tibialis$Spaceflight %in% c("Space_Flight"), ]
space_flight_t480<-space_flight$X480
sd(space_flight_t480) 

ground_control<-tibialis[tibialis$Spaceflight %in% c("Ground_Control"), ]
ground_control_t480<-ground_control$X480
sd(ground_control_t480) # has bigger variance, so will go in numerator

var.test(ground_control_t480,space_flight_t480)

```

## Question 3.2

Second, Use the two-sample t-test with either equal or unequal variances (based on your result above) to compare the means of Ca2+ ratios in the tibialis at times 20 and 480. Is there statistical support for a difference in Ca2+ ratios after space flight?\
- there is statistical support for a difference in Ca2+ ratios after space flight

```{r q3.2}

t20_ttest<-t.test(ground_control_t20,space_flight_t20, var.equal = T)
t480_ttest<-t.test(ground_control_t480,space_flight_t480, var.equal = T)
print(t20_ttest) # difference in mean is consistent with 0
print(t480_ttest) # difference in mean is not equal to 0

```

# T-Test - Two Sample Soleus

## Question 4.1

First, use the F-Test to determine if the variances of the two samples at time points 20 and 480 are statistically different from 1.\
- the sample variances at X20 and X480 are not significantly different from each other

```{r q4.1}

soleus<-ratio_data[ratio_data$Tissue %in% c("soleus_muscle"), ]

# Space Flight v Ground Control at X20
space_flight<-soleus[soleus$Spaceflight %in% c("Space_Flight"), ]
space_flight_s20<-space_flight$X20
sd(space_flight_s20) # has bigger variance, so will go in numerator

ground_control<-soleus[soleus$Spaceflight %in% c("Ground_Control"), ]
ground_control_s20<-ground_control$X20
sd(ground_control_s20) 

var.test(space_flight_s20,ground_control_s20)

# Space Flight v Ground Control at X480
space_flight<-soleus[soleus$Spaceflight %in% c("Space_Flight"), ]
space_flight_s480<-space_flight$X480
sd(space_flight_s480) 

ground_control<-soleus[soleus$Spaceflight %in% c("Ground_Control"), ]
ground_control_s480<-ground_control$X480
sd(ground_control_s480) # has bigger variance, so will go in numerator

var.test(ground_control_s480,space_flight_s480)

```

## Question 4.2

Second, Use the two-sample t-test with either equal or unequal variances (based on your result above) to compare the means of Ca2+ ratios in the tibialis at times 20 and 480. Is there statistical support for a difference in Ca2+ ratios after space flight?

-   there is not statistical support for a difference in Ca2+ ratios after space flight

```{r q4.2}

s20_ttest<-t.test(ground_control_s20,space_flight_s20, var.equal = T)
s480_ttest<-t.test(ground_control_s480,space_flight_s480, var.equal = T)
print(s20_ttest) # difference in mean is consistent with 0
print(s480_ttest) # difference in mean is consistent with 0

```

# Chi-Squared Test

## Question 5.1

What is the p-value of your test?\
- p-value = 0.533 - Not sure if this was a good test to do for this data because there are 0 Liquid Nitrogens above 1 - probably would be better to do the tibialis data or a different test

```{r q5.1}


joined<-full_join(ratio_data, full_data, join_by(Sample.Name == Sample.Name))
soleus<-joined[joined$Tissue %in% c("soleus_muscle"), ]

control<-soleus[soleus$Spaceflight %in% c("Cohort_Control","Ground_Control","Vivarium_Control"), ]

chi_data<-control[, names(control) %in% c("X480","Parameter.Value..Sample.Preservation.Method")]
rna<-chi_data[chi_data$Parameter.Value..Sample.Preservation.Method %in% c("RNALater"),]
liquid_nitrogen<-chi_data[chi_data$Parameter.Value..Sample.Preservation.Method %in% c("Liquid Nitrogen"),]

rna_g1<-as.numeric(sum(rna$X480 >= 1))
rna_l1<-as.numeric(sum(rna$X480 < 1))
ln_g1<-as.numeric(sum(liquid_nitrogen$X480 >= 1))
ln_l1<-as.numeric(sum(liquid_nitrogen$X480 < 1))

df<-data.frame("RNALater" = c(rna_g1,rna_l1), "Liquid Nitrogen" = c(ln_g1,ln_l1))
row.names(df)<- c(">= 1","< 1")

chisq.test(df)

```

## Question 5.2

Based on the results of this test (including any errors) what would you tell your labmate?\
- based on this test I would tell my labmate there is not evidence to suggest that the preservation method is not independent of the Ca2+ ratios. We would need more samples to support their hypothesis and would benefit from including more data.

# Additional hypotheses & Question 6

There are many more hypotheses we could test with this data. Now you need to come up with a hypothesis to test! Complete the steps below (1 point each)

1.  Write out a hypothesis/question to test.\

-   I hypothesize that there is one outlier data point in the Ca2+ ratios at time point X480.

2.  Select the appropriate test and provide the null and alternative hypotheses.\

-   I will use the Grubbs test.
-   null hypothesis: the data contains no outliers
-   alternative hypothesis: the data contains one outlier

3.  Conduct the test and report the associated p-value\

-   p-value = 1

```{r grubbs test}

grubbs_data<-ratio_data$X480
grubbs.test(grubbs_data)

```

4.  In 1-2 sentences, summarize your results\

-   These results do not indicate that the lowest value 0.5313 is an outlier. We fail to reject the null hypothesis that there are no outliers.
