---
title: "problem_set_10"
author: "Lydia Holley"
date: "2024-03-21"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(markovchain)
library(HMM)
```

## Problem A - Construct the Transmission and Emission matrices.

```{r problemA}

transmission<-matrix(data=c(0.9,0,0,1,0.1,0,0,0,0,1,0.8,0,0,0,0.2,0), nrow=4, byrow=F)

emission<-matrix(data=c((1/64),0.99,(1/61),0,(60/64),0.01,(60/61),0,(3/64),0,0,1), nrow=4, byrow=F)

colnames(transmission)<-c("intergenic","start","exon","stop")
rownames(transmission)<-c("intergenic","start","exon","stop")

colnames(emission)<-c("AUG","other","stop")
rownames(emission)<-c("intergenic","start","exon","stop")

transmission
emission
```
### Question A1 - Report your transmission matrix\
- in the code block

### Question A2 - Report your emission matrix\
- in the code block


## Problem B - Probability of our observation
```{r problemB}
# ACU AAA AUG AUG GCG UGA UUU

states<-c("intergenic","start","exon","stop")
obvs<-c("AUG","other","stop")

test_states<-c("intergenic","intergenic","start","exon","exon","stop","intergenic")
test_obvs<-c("other","other","AUG","AUG","other","stop","other")

mc<-new("markovchain", transitionMatrix=transmission)

pi<-steadyStates(mc)

hmm<-initHMM(States = states, Symbols = obvs, startProbs = c(1,0,0,0), transProbs = transmission, emissionProbs = emission)

log_probs<-forward(hmm,test_obvs)

probs<-exp(log_probs)

sum(probs[,7])

```
### Question B1 - Report the probability of the 7 states listed above when starting at an intergenic state\
- 0.0003972414



## Problem C - Hidden States
```{r problemC}

back_probs<-exp(backward(hmm, test_obvs))

posterior(hmm, test_obvs)

viterbi(hmm, test_obvs)

```

### Question C1 - What state has the maximum probability in the steady states?\
- Intergenic has the biggest probability in the steady states

### Question C2 - What is the probability associated with the steady-state selected in Question C1?\
- 0.5882353

### Question C3 - What is the most likely series of hidden states associated with the above DNA/RNA sequence (the 7 observed states from Question B)\
- intergenic, intergenic, intergenic, start, exon, stop, intergenic


## Problem D - Tuning a parameter (4 points)
```{r problemD}

pos_t<-c(0:100)/100

all_prob<-c()

for(i in 1:101){
 inter_to_inter<-pos_t[i]
 inter_to_start<-1-pos_t[i]
 A_this<-matrix(data=c(inter_to_inter,0,0,1,inter_to_start,0,0,0,0,1,0.8,0,0,0,0.2,0), nrow=4, byrow=F)
 mc_this<-new("markovchain", transitionMatrix=A_this)
 pi_this<-steadyStates(mc_this)
 hmm_this<-initHMM(States = states, Symbols = obvs, startProbs = pi_this, transProbs = A_this, emissionProbs = emission)
 this_p<-sum(exp(forward(hmm_this, test_obvs))[,7])
 all_prob<-append(all_prob, this_p)
}

plot(pos_t,all_prob)

```

### Question D1.1 - Plot the transition probabilities against the probability of our observation. The observation is the 7 states listed in question B.\
- included in code block

### Question D1.2 - Include the plot\
- included in code block

### Question D2 - What Intergenic transition probabilities maximize the probability of our observed sequence?\
- probabilities around 0.8 maximize the probability of our observed sequence
