---
title: "problem_set_9"
author: "Lydia Holley"
date: "2024-03-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library('expm')
```

# Problem_Set_9

## Problem A
```{r problemA}

prob <- matrix(ncol = 4, nrow = 4)
x <- c("to A", "to C", "to T", "to G")
y <- c("from A", "from C", "from T", "from G")
colnames(prob) <- y
rownames(prob) = x

prob[1,] <- c(0,0.333,0.333,0.333)
prob[2,] <- c(0.333,0,0.333,0.333)
prob[3,] <- c(0.333,0.333,0,0.333)
prob[4,] <- c(0.333,0.333,0.333,0)

# qA1
print(prob)

# qA2
df <- matrix(data=c(0,0,1,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")
                      
(prob %^% 4) %*% df

# qA3
q3_stop = FALSE
count = 0
while (q3_stop == FALSE){
  after<-format(round(((prob %^% count) %*% df),2))
  if (after[1] == 0.25 & after[2] == 0.25 & after[3] == 0.25 & after[4] == 0.25){
    print(count)
    print(after)
    q3_stop = TRUE
  }
  count <- count + 1
}

# qA4
df <- matrix(data=c(0.5,0,0,0.5),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")
                      
after4<-(prob %^% 4) %*% df
qa4<-after4[1]+after4[4]
print(qa4)
```

### Question A.1 - Report your transition matrix\
- in the code block

### Question A.2 - If you start at the base T, what is the probability that the state will return to T at 4 mutational events. Use the matrix math for this, which may require the expm package.\
- 0.2582238

### Question A.3 - After how many mutational events is the probability of being at any base equal if you start at state T. I.e. How many mutational events occur such that p(A)=p(T)=p(C)=p(G)=approximately 0.25. You should identify the number of steps at which R reports 0.25 as the value.\
- 5

### Question A.4 - If you start at either purine (A and G) with an equal probability, what is the probability that you will be at a purine after 4 mutational events?\
- 0.5041512


## Problem B

```{r problemB}
# qb1
prob2 <- matrix(ncol = 4, nrow = 4)
x <- c("to A", "to C", "to T", "to G")
y <- c("from A", "from C", "from T", "from G")
colnames(prob2) <- y
rownames(prob2) = x

probA<-9.4+10.4+27.2
probC<-9.4+35.4+8.4
probT<-35.4+10.4+9.2
probG<-9.2+8.4+27.2

prob2[1,] <- c(0,(9.4/probC),(10.4/probT),(27.2/probG))
prob2[2,] <- c((9.4/probA),0,(35.4/probT),(8.4/probG))
prob2[3,] <- c((10.4/probA),(35.4/probC),0,(9.2/probG))
prob2[4,] <- c((27.2/probA),(8.4/probC),(9.2/probT),0)

prob2

# qb2
df <- matrix(data=c(0,0,1,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")
                      
(prob2 %^% 4) %*% df

# qb3
state_prob<-list()

df <- matrix(data=c(0.25,0.25,0.25,0.25),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")

for (i in 1:20){
  after3b<-((prob2 %^% i) %*% df)
  state_prob<-append(state_prob,after3b[3,1])
}

plot(1:20,state_prob)

# qb4
df <- matrix(data=c(0.5,0,0,0.5),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")
                      
after4b<-(prob2 %^% 4) %*% df
qb4<-after4b[1]+after4b[4]
print(qb4)
```

### Question B.1 - Report your new transition matrix.\
- in the code block

### Question B.2 - If you start at the base T, what is the probability that the state will return to T at 4 mutational events. Use the matrix math for this, which may require the expm package.\
- 0.3683680

### Question B.3 - Plot the probability of the state being T for 20 mutational changes.\
- in code block

### Question B.4 - If you start at either purine with an equal probability, what is the probability that you will be at a purine after 4 mutational events?\
- 0.4610521


## Problem C
```{r problemC}

# qc1
prob3 <- matrix(ncol = 4, nrow = 4)
x <- c("to A", "to C", "to T", "to G")
y <- c("from A", "from C", "from T", "from G")
colnames(prob3) <- y
rownames(prob3) = x

probA<-9.4+10.4+27.2
probC<-9.4+35.4+8.4
probT<-35.4+10.4+9.2
probG<-9.2+8.4+27.2

prob3[1,] <- c(999,(9.4/probC),(10.4/probT),(27.2/probG))
prob3[2,] <- c((9.4/probA),999,(35.4/probT),(8.4/probG))
prob3[3,] <- c((10.4/probA),(35.4/probC),999,(9.2/probG))
prob3[4,] <- c((27.2/probA),(8.4/probC),(9.2/probT),999)

prob3<-prob3/1000

prob3

# qc2
df <- matrix(data=c(0,0,1,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")
                      
after2c<-(prob3 %^% 4) %*% df

qc2<- after2c[1]+after2c[2]+after2c[4]

qc2

# qc3
df <- matrix(data=c(0,0,1,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")
                      
after3c<-(prob3 %^% 400) %*% df

qc3<- after3c[1]+after3c[2]+after3c[4]

qc3

# qc4

# A
df <- matrix(data=c(1,0,0,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")

(prob3 %^% 10000) %*% df

# C
df <- matrix(data=c(0,1,0,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")

(prob3 %^% 10000) %*% df

# T
df <- matrix(data=c(0,0,1,0),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")

(prob3 %^% 10000) %*% df

# G
df <- matrix(data=c(0,0,0,1),nrow=4,byrow=T)
colnames(df) <- "prob"
rownames(df) <- c("A", "C", "T", "G")

(prob3 %^% 10000) %*% df

````
### Question C.1 - Report your transition matrix\
- in the code block

### Question C.2 - If you start at the base T, what is the probability that the state will be any other state than T after 4 mutational events. Use the matrix math for this, which may require the expm package.\
- 0.003990983

### Question C.3 - If you start at the base T, what is the probability that the state will be any other state than T after 400 mutational events. Use the matrix math for this, which may require the expm package.\
- 0.3015315

### Question C.4 - After 10,000 mutational events, what base(s) are you most likely to end up for each of the starting bases:
- A: T
- T: T
- C: T
- G: T
