---
title: "problem_set_2"
author: "Lydia Holley"
date: "2024-01-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load needed libraries
library(naniar)
library(dplyr)
library(naniar)
library(stringr)
```

# Step 1 - Import the Data

The data will be imported here

```{r data import}
#load in data
sw1<-read.delim("star_wars_data.1.tsv",header = T, sep = "\t")
sw2<-read.delim("star_wars_data.2.txt",header = T, sep = ",")

```

# Step 2 - Clean up the data

Data will be cleaned up here

```{r clean data}
#find initial number
sum(is.na(sw1))

# replace unknowns with proper NA
sw1<-replace_with_na(sw1,replace = list(hair_color="unknown"))
sw1<-replace_with_na(sw1,replace = list(height="unknown"))
sw1<-replace_with_na(sw1,replace = list(mass="unknown"))
sw1<-replace_with_na(sw1,replace = list(skin_color="unknown"))
sw1<-replace_with_na(sw1,replace = list(eye_color="unknown"))
sw1<-replace_with_na(sw1,replace = list(birth_year="unknown"))
sw1<-replace_with_na(sw1,replace = list(sex="unknown"))
sw1<-replace_with_na(sw1,replace = list(gender="unknown"))
sw1<-replace_with_na(sw1,replace = list(homeworld_species="unknown"))

# check final NA
sum(is.na(sw1))

```

## **Question 1 - How many \_NA_s are in the cleaned dataset**

There are 95 NAs in the cleaned dataset

# Step 3 - Clean up the last column

Split the homeworld_species column into 2 columns.

```{r split column}
# get homeworld and species columns
sw1$homeworld<-str_split_i(sw1$homeworld_species, "_", 1)
sw1$species<-str_split_i(sw1$homeworld_species, "_", 2)

# remove combined column
sw1<-subset(sw1, select = -homeworld_species)

# find answer to question 2 by indexing and counting rows
de_sp <- sw1[sw1$species %in% c('Droid', 'Ewok'), ]
nrow(de_sp)
```

## **Question 2 - How many species are either a Droid or an Ewok**

There are 7 Droid or Ewok species.

```{r replace NA}
# replace improper NA with proper NAs
sw1<-replace_with_na(sw1,replace = list(homeworld="NA"))
sw1<-replace_with_na(sw1,replace = list(species="NA"))

# find number of homeworlds ignorings NAs
n_distinct(sw1$homeworld,na.rm = T)

```

## **Question 3 - How many different home worlds are there? Do not include missing data**

There are 48 different home worlds

# Step 4 - Edit the mass column

*already fixed the unknowns in this column*

```{r median}
# find the median mass, ignoring NAs
median(as.numeric(sw1$mass), na.rm = T)

```

## **Question 4 - What is the median mass of all the characters?**

The median mass of all the characters is 79.

# Step 5 - Edit star_wars_data.2.txt

This chunk will create columns for the film number

```{r film}
#split column
sw2$first<-str_split_i(sw2$films, "AND", 1)
sw2$second<-str_split_i(sw2$films, "AND", 2)
sw2$third<-str_split_i(sw2$films, "AND", 3)
sw2$fourth<-str_split_i(sw2$films, "AND", 4)
sw2$fifth<-str_split_i(sw2$films, "AND", 5)
sw2$sixth<-str_split_i(sw2$films, "AND", 6)
sw2$seventh<-str_split_i(sw2$films, "AND", 7)

# remove film column
sw2<-subset(sw2, select = -films)

#answer question 5 by indexing and counting rows
sec_film <- sw2[sw2$second %in% c(' Return of the Jedi',' Return of the Jedi '), ]
nrow(sec_film)
```

## **Question 5 - How many character's 2nd film was "Return of the Jedi"?**

5 character's 2nd film was "Return of the Jedi"

```{r one film}
# how many characters have NA in the "second" column
sum(is.na(sw2$second))

```

## **Question 6 - How many characters were only in 1 film?**

46 characters were only in 1 film

# Step 6 - Merge the datasets

Combine the two datasets to answer the questions below

```{r combine datasets}
# join datasets
joined<-full_join(sw1, sw2, join_by(name == names))

# subset groups
one_movie<- joined[ which(is.na(joined$second)),]
more_movies<- joined[ which(!is.na(joined$second)),]

# find means of subset
mean(one_movie$height, na.rm = T)
mean(more_movies$height, na.rm = T)

```

## **Question 7 - Which group has a greater mean height - characters in 1 movie or characters in more than 1 movie?**

Characters in 1 movie have a greater mean height, but only by a very small amount (174.95 v 174.2683)

```{r empire}
# subset the 2nd movie criteria
empire2<-joined[joined$second %in% c(' The Empire Strikes Back',' The Empire Strikes Back '), ]

# subset human criteria
emp_hum<-empire2[empire2$species %in% c('Human'),]

# find answer to question 8
nrow(emp_hum)
```

## **Question 8 - How many Human characters had their 2nd movie "The Empire Strikes Back"?**

6 human characters had their 2nd movie "The Empire Strikes Back".

```{r ratio}
# only keep characters that appear in a second movie
# we already have this from earlier, it is more_movies

# subset into early movies
early<-more_movies[more_movies$second %in% c(' The Empire Strikes Back',' The Empire Strikes Back ', ' Return of the Jedi',' Return of the Jedi ', ' A New Hope',' A New Hope '), ]

# subset into later movies
later<-more_movies[!(more_movies$second %in% c(' The Empire Strikes Back',' The Empire Strikes Back ', ' Return of the Jedi',' Return of the Jedi ', ' A New Hope',' A New Hope ')), ]

#find early ratio
emass<-median(as.numeric(early$mass), na.rm = T)
eheight<-median(as.numeric(early$height), na.rm = T)
emass/eheight

#find later ratio
lmass<-median(as.numeric(later$mass), na.rm = T)
lheight<-median(as.numeric(later$height), na.rm = T)
lmass/lheight
```

## **Question 9 - You hypothesize that the ratio of mass/height is smaller in characters from the earlier movies. Which group has a smaller median mass/height ratio - characters whose second movie was in the original trilogy ("The Empire Strikes Back," "Return of the Jedi," "A New Hope") or those who had a second movie from the later series?**

Characters who had a second movie from the later series had a smaller median mass/height ratio.
