---
title: "Problem_Set_1"
author: "Lydia Holley"
output: pdf_document
date: "2024-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)
```

# Step 3 - Import the Data

*Data will be imported here using read.table*

```{r data import}
data <- read.table("GSE164805_series_matrix.txt", header = TRUE)

```

## Find Solutions for Questions 2-7

*This chunk contains the code for questions 2-7*

```{r solutions2_7}
# find rows and columns
rows = nrow(data)
cols = ncol(data)

# find how data is stored
str(data)

# determine if there are null values in the dataset
# is.na() returns true when there are null values
# sum of that command give total number of true returns
# <https://www.geeksforgeeks.org/how-to-find-and-count-missing-values-in-r-dataframe/>
# used above page as reference
q5 = sum(is.na(data))

# check if column is a factor (given in assignment)
q6 = is.factor(data$ID_REF)

# for question 7 using as.factor to change data type and $ to get column

```

**Question 2 - *How many rows are in the dataset?***\
There are 60795 rows in the dataset.

**Question 3 - *How many columns are in the dataset?***\
There are 16 columns in the dataset.

**Question 4 - *Is the object storing our data a data frame or a matrix?***\
The object storing our data is a data frame.

**Question 5 - *Are there any NAs in our dataset?***\
There are not any NAs in our data.

**Question 6 - *Is the column ID_REF a factor?***\
No, the column ID_REF is not a factor.

**Question 7 - *What is the line of code you would use to set the column ID_REF as a factor?***\
The line of code I would use is: `data$ID_REF <- as.factor(data$ID_REF)`

# Step 4 - Subset the data frame

*This chunk will subset the data to include only the desired data*

The genes (ID_REFs) ASHG19AP1B100028737V5 ASHG19AP1B100017453V5 ASHG19AP1B100004705V5

For the individuals (columns) GSM5019817 GSM5019827 GSM5019828

```{r subset}

# get desired columns
s_data <- data[,c('ID_REF','GSM5019817', 'GSM5019827', 'GSM5019828')]

# get desired rows
s_data <- subset(s_data, ID_REF %in% 
          c('ASHG19AP1B100028737V5','ASHG19AP1B100017453V5', 'ASHG19AP1B100004705V5'))

# find mean for GSM5019817 columm
q8 <- mean(s_data[['GSM5019817']])

```

**Question 8 - *What is the mean expression across the three genes for GSM5019817?***\
The mean expression across the three genes for GSM5019817 is 3.95154.

# Step 6 - Change the subset dataset to long format

*This chunk will change the dataset to long format*

```{r change subset}

long_data <- melt(s_data, id.vars=c("ID_REF"), variable.name="Individuals", value.name="Expression")

long_data
```

**Question 9 - *How many observations in our subset data have a gene expression value greater than 4?***\
*This chunk will determine number of expressions \> 4*

```{r q9}
Q9 <- sum(long_data$Expression > 4)

```

There are 6 observations in our subset data with a gene expression value greater than 4.
